Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex

\begin{equation}
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
\end{equation}

@@

The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators. The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.  Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.  The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code. We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.

@@

\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
this using induction \cite{bal2005convergence}.

@@

% I don't understand the above part ~ Andy

@@

From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.  Of this, the only parallelizable part is the $kQt$ term.  Thus to find the expression of the parallelizable portion, we simplify:

\[
AA
\]

@@

\[
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\[
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\begin{center}
\begin{tabular}{|c|c|} \hline
($N_x, N_y$) & Time (seconds)\\ \hline
(128, 128) & 0.987\\ \hline
(256, 256) & 7.889\\ \hline
(512, 512) & 63.913\\ \hline
\end{tabular}
\end{center}

@@

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
(128, 128) & 2 & 2 &  \\ \hline
(128, 128) & 4 & 2 & 2.31 \\ \hline
(128, 128) & 4 & 4 &  \\ \hline
(128, 128) & 8 & 8 &  \\ \hline
(256, 256) & 2 & 2 & 2.57 \\ \hline
(256, 256) & 4 & 2 & 11.32 \\ \hline
(256, 256) & 4 & 4 & 13.65 \\ \hline
(256, 256) & 8 & 8 & 17.44 \\ \hline
(512, 512) & 2 & 2 & \\ \hline
(512, 512) & 4 & 2 &  \\ \hline
(512, 512) & 4 & 4 &  \\ \hline
(512, 512) & 8 & 8 &  \\ \hline
\end{tabular}
\end{center}

Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, linear scaling is achieved. However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed). Hence, this is an example of weak scaling. However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes. Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.

\subsection{Parallelism Paradigm Comparisons}

Now with 3 different parallel implementations, two in time with the parareal, one in space, we compare across the methods to see if we can make any statements about general performance of our different paradigms.

TODO WESLEY

TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand

TODO bsim and ashi show graphs and some times for the forward Euler parareal
versions---the fastest ones or both if we have both implemented

\subsection{Comparison to C++ Implementations}

The numpy environment combined with the required structure of MPI required
certain type/data structure conversions which will take time.not not cunot cu

@@

\subsection{Demonstrating Scaling}

ashi TODO by taking some numbers and following

@@

@@

TODO Wesley

>>>>

\begin{equation}
S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
\end{equation}

@@

The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators. The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.  Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.  The exact optimum pairing is problem-specific and the code is written so that a different ``n-order method step'' function could be written and then replaced into the code. We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.

@@

\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
this using induction~\cite{bal2005convergence}.

@@

@@

From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + k(t+ Qt)$.  Of this, the only parallelizable part is the $kQt$ term.  Thus to find the expression of the non-parallelizable portion, we simplify:

\[
B = \frac{t+kt}{t+kt+kQt} = \frac{1+k}{kQ+k+1}
\]

@@

\[
\begin{aligned}
S(n) &= \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{1}{\left( B + \frac{1}{n} (1-B) \right)}\\
       &= \frac{T(1)}{T(1) \left( \frac{1+k}{kQ+k+1} + \frac{1}{n} (1- \frac{1+k}{kQ+k+1}) \right)} \\
       &= \frac{n(kQ+k+1)}{n+nk+kQ}
\end{aligned}
\]

As written below, we have used $Q = 100$ with various $k$.

When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of $S(n) = \frac{203n}{3n+200} \approx 67$

When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of $S(n) = \frac{1011n}{11n+1000} \approx 91$

It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel. This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds. We do see however large constants in the denominator which means that we will get better speedup as we increase n past a certain size.

@@

\[
S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\begin{center}
\begin{tabular}{|c|c|} \hline
($N_x, N_y$) & Time (seconds)\\ \hline
(128, 128) & 0.987\\ \hline
(256, 256) & 7.889\\ \hline %real
(512, 512) & 63.913\\ \hline %real
\end{tabular}
\end{center}

@@

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
(128, 128) & 4 & 2 & 2.31 \\ \hline
(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
(256, 256) & 2 & 2 & 2.57 \\ \hline %real
(256, 256) & 4 & 2 & 11.32 \\ \hline
(256, 256) & 4 & 4 & 13.65 \\ \hline
(256, 256) & 8 & 8 & 17.44 \\ \hline
(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
\end{tabular}
\end{center}

Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved. However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed). Hence, this is an example of weak scaling. However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes. Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.

@@

\subsection{Demonstrating Scaling and Efficiency}

\subsection{Parallel in Space}

bsim TODO eventually

@@

TODO NUMPY WRITE FATST HING

@@

TODO WRITE ABOUT BETTER DATA COLLECTION
